Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-03-28 figures/neural networks/Single_CNN.tex
Updates from Overleaf

		\draw[fill=black,opacity=0.2,draw=black] (5.5,1.5) -- (7.5,1.5) -- (7.5,3.5) -- (5.5,3.5) -- (5.5,1.5);
		\draw[fill=black,opacity=0.2,draw=black] (5,1) -- (7,1) -- (7,3) -- (5,3) -- (5,1);
		\draw[fill=black,opacity=0.2,draw=black] (4.5,0.5) -- (6.5,0.5) -- (6.5,2.5) -- (4.5,2.5) -- (4.5,0.5);
		\draw[fill=black,opacity=0.2,draw=black] (4,0) -- (6,0) -- (6,2) -- (4,2) -- (4,0);
	\end{tikzpicture}
	\caption[Illustration of a convolutional layer.]{Illustration of a single convolutional layer. If layer $l$ is a convolutional layer, the input data (if $l = 1$) or a feature map of the previous layer is convolved by different filters to yield the output feature maps of layer $l$.}
	\label{fig:convolutional-layer}
\end{SCfigure}

>>>>

		\draw[fill=black,opacity=0.2,draw=black] (5.5,1.5) -- (7.5,1.5) -- (7.5,3.5) -- (5.5,3.5) -- (5.5,1.5);
		\draw[fill=black,opacity=0.2,draw=black] (5,1) -- (7,1) -- (7,3) -- (5,3) -- (5,1);
		\draw[fill=black,opacity=0.2,draw=black] (4.5,0.5) -- (6.5,0.5) -- (6.5,2.5) -- (4.5,2.5) -- (4.5,0.5);
		\draw[fill=black,opacity=0.2,draw=black] (4,0) -- (6,0) -- (6,2) -- (4,2) -- (4,0);
	\end{tikzpicture}
	\caption[Illustration of a 2D convolutional layer.]{Illustration of a 2D single convolutional layer. If layer $l$ is a convolutional layer, the input data (if $l = 1$) or a feature map of the previous layer is convolved by different filters (or kernels) to yield the output feature maps of layer $l$.}
	\label{fig:convolutional-layer}
\end{SCfigure}

### 2023-03-28 introduction/Subsections/Motivation.tex
Updates from Overleaf

Let's pick a core MSA task as a sample to display such complexity: Music Structure Analysis (MSA) is an interdisciplinary field that aims to understand the structure of music. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification. The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity.

@@

\begin{figure}[!ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. We can see the passage's original score, the schenkerian unfolding of the melody, the chord degrees, and the tonal function.}
\label{fig:timeseries}
\end{figure}

>>>>

Let's pick a core MIR task as a sample to display such complexity: Music Structure Analysis (MSA), an interdisciplinary field that aims to understand the structure of music\cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification. The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity.

@@

\begin{figure}[!ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. We can see the passage's original score, the schenkerian unfolding of the melody, the chord degrees, and the tonal function.}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

### 2023-03-28 introduction/Subsections/Objectives.tex
Updates from Overleaf

In recent years, using neural networks has enabled low-dimensional musical embeddings that capture crucial musical features and allow the computation of content-based similarity between musical passages. By acquiring an embedding function $f(x) \in \mathbb{R}^d$ that maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, we can measure the similarity between two musical passages using the Euclidean distance, as represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The effectiveness of such systems depends on the embedding space's ability to interpret music in a way that aligns with human perception. Therefore, a suitable embedding space is crucial to ensure the output aligns with human expectations.

>>>>

In recent years, using neural networks has enabled low-dimensional musical embeddings that capture crucial musical features and allow the computation of content-based similarity between musical passages. By acquiring an embedding function $f(x) \in \mathbb{R}^d$ that maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, we can measure the similarity between two musical passages using the Euclidean distance, as represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The effectiveness of such systems depends on the level of musical abstraction and the embedding space's ability to interpret music in a way that aligns with human perception. Therefore, a suitable embedding space is crucial to ensure the output aligns with human expectations.

### 2023-03-28 methods/subsections/machine learning.tex
Updates from Overleaf

Machine Learning (ML), particularly Deep Learning, has significantly advanced the Music Information Retrieval (MIR) field by enabling automatic feature learning, improving classification and analysis tasks, and facilitating novel applications like music generation and source separation. Adopting techniques like Convolutional Neural Networks (CNNs), Recurrent Neural Networks (RNNs), and generative models has led to better performance in tasks such as music classification, tagging, recommendation, transcription, synthesis, and analysis. As a result, ML has unlocked the potential of MIR, revolutionizing how we interact with, understand, and create music.

@@

@@

%Single neuron
\input{figures/neural networks/Single Neuron}
%NN
\input{figures/neural networks/Network_graph}
%Multi-layer_NN
\input{figures/neural networks/Multi-layer_NN}
%Backpropagation
\input{figures/neural networks/Back_propagation}

\section{Convolutional Neural Network (CNN)}

@@

\item Convolutional layers: These layers apply a convolution operation to the input data, effectively learning local patterns and features in the image. The raw audio data is typically transformed into a spectrogram or a Mel-spectrogram representation, which can be considered an image-like 2D representation of the audio data. The convolutional layers are designed to learn local patterns and features in the audio data
\vspace*{3mm}

\item Pooling layers: These layers downsample the data, reducing the spatial dimensions of the input while retaining important information. They can also be used to downsample the data and reduce the temporal dimensions of the input while retaining important information. 
\vspace*{3mm}

\item Fully connected layers: These layers connect every neuron in one layer to every neuron in another, allowing the network to learn non-linear combinations of the features learned in the previous layers.
\vspace*{3mm}

\item Output layer: The output layer produces the final predictions of the network.
\end{itemize}

@@

%CNN
A high-level illustration of a simple convolutional neural network indicating convolutional layers, pooling layers, and fully-connected layers without details on the number of channels or neurons per layer or the input data size can be seen in XXX.
\input{figures/neural networks/CNN_diagram}

>>>>

\section{Introduction}
Machine Learning (ML), and particularly Deep Learning, has significantly advanced the MIR field by enabling automatic feature learning, improving classification and analysis tasks, and facilitating novel applications like music generation and source separation. Adopting musically motivated neural network architectures \cite{musicmotivCNN} blends together musical concepts and perceptual approaches and use this knowledge to analyze and understand music in a way that goes beyond traditional music analysis techniques.

@@

%Single neuron
\input{figures/neural networks/Single Neuron}

@@

%NN
\input{figures/neural networks/Network_graph}

%Multi-layer_NN
\input{figures/neural networks/Multi-layer_NN}

\section{Deep Learning (DL)}

%Backpropagation
\input{figures/neural networks/Back_propagation}

\subsection{Convolutional Neural Network (CNN)}

@@

\item \textbf{Convolutional layers:} These layers apply a convolution operation to the input data. The raw audio data can be either a 1D signal or transformed into a spectrogram or a Mel-spectrogram representation, which can be considered an image-like 2D representation of the audio data. The convolutional layers are designed to effectively learn local patterns and features in the audio data \footnote{For a visual explanation of convolution, see this video: \url{https://youtu.be/KuXjwB4LzSA}}.

\textbf{1D Convolution Example:} Given a 1D input signal $I = [i_1, i_2, ..., i_n]$ and a filter (or kernel) $K = [k_1, k_2, ..., k_m]$, the convolution operation is defined as $(I * K)j = \sum{u=1}^{m} I_{j-u+1} K_u$, where $j$ ranges from $1$ to $n-m+1$, and $I * K$ represents the resulting feature vector after applying the convolution. The objective of the 1D convolution operation is to slide the filter over the input signal and compute the element-wise product and sum for each local region, effectively learning and detecting local patterns in the input data.

\item \textbf{Pooling layers:} These layers downsample the data, reducing the spatial or temporal dimensions of the input while retaining important information. A common pooling operation is the max-pooling, which can be defined as follows: given an input matrix $I$ and a pooling window of size $m \times n$, the max-pooling operation selects the maximum value within each non-overlapping $m \times n$ region of the input matrix. This operation effectively reduces the input's dimensions while retaining the most important information in each region.
\vspace*{3mm}

\item \textbf{Fully connected layers:} These layers connect every neuron in one layer to every neuron in another, allowing the network to learn non-linear combinations of the features learned in the previous layers. Mathematically, a fully connected layer can be represented as $Y = XW + b$, where $X$ is the input matrix, $W$ is the weight matrix, $b$ is the bias vector, and $Y$ is the output matrix. The non-linearity is introduced through the activation function, which is applied element-wise to the output matrix, such as $Y' = f(Y)$, where $f$ represents the activation function and $Y'$ is the transformed output matrix.
\vspace*{3mm}

\item \textbf{Output layer:} The output layer produces the final predictions of the network.

\end{itemize}

@@

%CNN
\input{figures/neural networks/CNN_diagram}

### 2023-03-28 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

\section{Model architecture and training strategy}
Self-Supervised Learning (SSL) is a technique to learn latent representations from large-scale data without relying on human annotations. Instead, it solves pretext tasks by creating different views of an object that are highly correlated. An SSL model is then trained to generalize objects' representations in a latent high-dimensional space. To achieve this, the model contrasts the representations of the same object with other objects (defined as negative samples) during training. By doing so, a contrastive SSL model is expected to produce more distinctive representations.

@@

We propose using a Triple Siamese Network to train a model for music similarity retrieval. Specifically, we aim to minimize the loss function between the anchor, positive, and negative samples by using online triplet mining.

>>>>

\section{Model architecture and training strategy}
\section{Self-Supervised Learning (SSL)}
Self-Supervised Learning (SSL) is a technique that leverages inductive bias to learn latent representations from large-scale data without relying on human annotations. By focusing on pretext tasks that create different, highly correlated views of an object, SSL models exploit the underlying structure and known variance in the data.

An SSL model is trained to generalize objects' representations in a latent high-dimensional space. To achieve this, the model contrasts the representations of the same object with other objects (defined as negative samples) during training. By doing so, a contrastive SSL model is expected to produce more distinctive representations \cite{audioselfsupsurvey}. Several MIR fields, such as music genre classification, or music similarity (both in the symbolic \cite{musicviaselfsup} and time domain \cite{epidemic}), have taken advantage of SSL techniques to improve their performance and alleviate the need for large amounts of labeled data \cite{audioselfsupsurvey}.

\subsubsection{Strengths}

\begin{enumerate}
    \item Reduces dependency on labeled data: SSL capitalizes on inductive bias to learn from large amounts of unlabeled data, which is particularly beneficial when obtaining labeled data is expensive or time-consuming.
    \item Encourages feature extraction: SSL encourages learning useful features or representations from the data, which can lead to better generalization and improved performance on downstream tasks.
    \item Transfer learning: SSL models can be fine-tuned for specific tasks with smaller labeled datasets, allowing for more efficient training and potentially better results.
    \item Robustness: SSL models can learn robust features from noisy or incomplete data, improving their ability to handle real-world scenarios.
\end{enumerate}

\subsubsection{Flaws}

\begin{enumerate}
    \item Less interpretable: SSL models tend to be less interpretable than supervised models, as their learning mechanisms are less explicit and rely more on inductive bias.
    \item Computational complexity: SSL models often require more training iterations and larger batch sizes, which can result in higher computational costs.
    \item Sensitive to hyperparameters: SSL models can be sensitive to hyperparameters such as learning rate, batch size, and the choice of the contrastive loss function, which may require extensive tuning.
    \item Limited applicability: SSL may not be suitable for all tasks or domains, as its performance depends on the quality and relevance of the self-supervised pretext tasks and the effectiveness of the inductive bias in capturing the known variance in the data.
\end{enumerate}

\subsection{Contrastive Learning of Musical Representations (CLMR)}

Contrastive Learning of Musical Representations (CLMR) is a self-supervised learning technique that aims to learn discriminative representations of music by leveraging its structure and content without explicit labels \cite{CLMR2021}\cite{multisourceclmr2023}. The learning process involves creating different augmentations or transformations of the same musical piece as positive pairs and using other randomly selected pieces as antagonistic pairs. A neural network is then trained to minimize a contrastive loss function, which encourages the model to produce similar representations for positive pairs and dissimilar representations for opposing pairs. Once trained, these learned representations can be utilized for various MIR downstream tasks.

@@

We propose using a Triple Siamese Network to train a model for music similarity retrieval \cite{contentmusicsimtriplet2020}. Specifically, we aim to minimize the loss function between the anchor, positive, and negative samples by using online triplet mining.

### 2023-03-28 preface/preface.tex
5-2-2023 commit
Updates from Overleaf

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{preface}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
(Optional, if used placed on a right page next to an empty left page)

I would like to dedicate this work to...
\newpage

\end{preface}

>>>>

\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{preface}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)

Music and science have always been my two greatest passions. When the opportunity to work at Epidemic Sound presented itself, it was as if the universe had conspired to bring my interests together, offering me the chance to explore the fascinating Music Information Retrieval (MIR) field. With the support and encouragement of my loved ones, I took a leap of faith and moved to Stockholm to embark on this exciting journey.

This work aims to help untangle the complexities of music and offer a more musically-driven perspective to MIR. I aim to contribute to a deeper understanding of the intricate relationships within musical data by bridging the gap between music theory and computational analysis.

I am deeply grateful to the Epidemic Sound team for their guidance, expertise, and camaraderie throughout this project. I would also like to extend my heartfelt appreciation to my family for their unwavering support and belief in my abilities.

My motivation for writing this piece stems from a desire to grow as a musician, programmer, professional, and individual. I believe that by delving into the world of MIR, I can expand my horizons while making a meaningful contribution to the field.

This work is intended for any curious human being, regardless of their background in music or computer science. I hope it will inspire others to explore the captivating intersection of these two disciplines.

The scope of this work encompasses a range of MIR tasks and methodologies, as well as discussions on the challenges and opportunities that arise within the field. Despite the ambitious nature of this project, I acknowledge that time and my academic background limitations may have constrained the depth of my analysis. Nevertheless, I hope this work will serve as a starting point for further exploration and inspire new ideas in the realm of MIR.

\newpage
\end{preface}
